


% ======================================================================
% == Institue and Project Period
% ======================================================================
\vspace*{\fill} % > Shift it to bottom

\thispagestyle{empty} % > Don't apply pagestyles to this page

\begin{center}
% \noindent
A digital version of this thesis will be made available at:\\
\texttt{https://github.com/markur4/Dissertation}

The \LaTeX~source code is found in this subfolder:\\
\texttt{./THESIS}

Videos are found in this subfolder:\\
\texttt{./publications/Chapter1 Supplementary/}

\vspace{\vdouble}

% \noindent
This work was conducted at the Department of Musculoskeletal Tissue
Regeneration (Bernhard-Heine-Centre for Locomotive Research), University
of Würzburg from 08.10.2018 to 
\DTMtoday~ under the supervision of Prof Dr.\,rer.\,nat.~Regina Ebert.
% 31.03.2024 
\end{center}




% ======================================================================
% == Acknowledgements
% ======================================================================

\newpage
\thispagestyle{empty} % > Don't apply pagestyles to this page

\section*{Acknowledgements} % > The * means it's not listed in TOC
\begin{center}%
    \small
    \noindent I am deeply grateful to the following supporters who have
    contributed to the completion of this PhD thesis in myriad ways:

    \vspace{\vhalf}

    \acknowledge{Doris Schneider}{%
        for her exceptional assistance in laboratory work. Her patience, focus, and
        dedication were invaluable to the experimental success of this
        research.%
    }

    \acknowledge{Regina Ebert}{%
        for her unwavering support and insightful guidance. Her patience and
        encouragement were instrumental in making \texttt{plotastic} a reality,
        and her feedback significantly improved the quality of this thesis.%
    }
    
    \acknowledge{Franziska Jundt \& Torsten Blunk}{%
    for their invaluable supervision of this project. Their professional
    feedback and honest critiques pushed me to strive for the highest quality in
    my work.%
}

    \acknowledge{Marie-Nicole Kobsar}{%
        my beloved significant other, for her endless patience, kindness, and
        support throughout the entire doctoral journey. Her encouragement and
        understanding were a source of strength and motivation.%
    }

    \acknowledge{My Lab Colleagues \& Collaborators}{%
        for their countless contributions, insights and constructive feedback
        throughout my doctorate. 
        % Special thanks to Melanie Krug, Jutta Meißner-Weigl, Sabine Zeck, Marietta
        % Herrman, Sigrid Müller-Deubert, Wyonna Rindt, Franz Jakob, Denitsa
        % Docheva, Drenka Trivanović, Ellen Leich, Franziska Jundt, Dirk Hose,
        % Tanja Nicole Hartmann, Torsten Blunk, and Maximilian Rudert.%
    }

    \acknowledge{My Family}{%
        for their unconditional support, encouragement, and feedback throughout
        my academic journey. Their belief in me provided the foundation upon
        which this achievement was built.%
    }

    \vspace{\vhalf}

    \noindent Additionally, I wish to extend my gratitude to all my friends who
 listened, discussed, and helped in various ways.

    \noindent I would also like to acknowledge the funding bodies and
    institutions that supported this research, including the German Research
    Foundation (DFG), the Graduate School of Life Sciences (GSLS), and the University of
    Würzburg, whose contributions made this work possible.

    \noindent Finally, a special thanks goes to the Elite Network Bavaria (ENB)
    and the University of Bayreuth for my education in molecular cell biology
    and biological physics. I felt well-prepared for my doctorate at the
    University of Würzburg, and am very grateful for the opportunities for
    interdisciplinary exchange with amazing physicists and other scientists.

    \noindent \textbf{Thank you all for your incredible support and encouragement.}

\end{center}