

\unnsubsection{\textit{\textbf{Aims}}}% > Aims are part of Introduction
\label{sec:aims}%
This PhD thesis aims to enhance our understanding of myeloma cell behavior and
improve the analysis of complex biomedical datasets. The specific aims are as
follows:

\begin{itemize}
    \item Develop an \textit{in vitro} model to elucidate the mechanisms of
          myeloma cell dissemination in interaction with mesenchymal stromal cells
          (hMSCs), focusing particularly on:
          \begin{itemize}
              \item Observing and quantifying cell proliferation, attachment, and
                    detachment dynamics using time-lapse microscopy.
              \item Isolating and characterizing distinct myeloma subpopulations
                    interacting with hMSCs to understand differential gene expression
                    related to cell adhesion and patient survival.
          \end{itemize}

    \item Design and implement a Python-based software tool, \texttt{plotastic},
          to facilitate the analysis of multidimensional datasets
          generated in biomedical research. This tool will aim to:
          \begin{itemize}
              \item Streamline the data analysis process, making it more efficient and
                    transparent.
              \item Integrate visualization and statistical analysis capabilities to
                    ensure that data analysis protocols are aligned with the ways in which
                    data is visualized.
              \item Provide a case study demonstrating the application of
                    \texttt{plotastic} in the analysis of \textit{in vitro} dissemination
                    experiments, discussing the tool's benefits in handling semi-big data and
                    potentially enhance reproducibility.
          \end{itemize}

    \item Synthesize the findings from the experimental and software development
          components to advance the understanding of myeloma dissemination and improve
          research practices in biomedical data analysis.
\end{itemize}

% These aims are crafted to address both the biological and technical challenges
% in current cancer research methodologies and data science applications in
% biomedicine, providing advancements that could help with designing novel therapeutic
% strategies and more robust scientific inquiries.
These aims are designed to tackle both the biological and technical challenges
in current cancer research methodologies and data science applications in
biomedicine, potentially contributing to the development of novel
anti-disseminative therapeutic strategies and more rigorous scientific
investigations.