




% 
% ======================================================================
\unnsubsection{How Exploratory Live-Cell Imaging Transformed the Research Focus}%
\label{sec:discussion_potential_breakthroughs}%
Exploratory experimentation emphasizes discovering and characterizing novel
phenomena \cite{mattigClassifyingExploratoryExperimentation2022}. Exploratory
cell biology often leverages emerging technologies to visualize and analyze the
mechanisms of cell behavior dynamically. Such approaches allow real-time
observations that can lead to unexpected insights and breakthroughs. In this
project, the application of live-cell imaging proved pivotal.

\textbf{Direct Observation of Complexity and Novelty:}
Initially, the project did not focus on \textit{in vitro} myeloma cell
dissemination. The project's research focus shifted when making the unexpected
\dashed{or argueably insignificant} observation of cancer cells
detaching from aggregates. This shows the transformative power of time-lapse
microscopy or live cell imaging \cite{coleLivecellImaging2014}. For the author,
live-cell imaging provides an observation method that's unmatched in intuition
and directness. Unlike RNA sequencing, which can obscure biological processes
behind cryptic data, live-cell imaging offers a clear view into the dynamic
cellular events as they unfold.

Such clarity was particularly effective in revealing the detachment of cells
following division, a phenomenon that might be overlooked in static analyses.
Multiple parameters can be read out in parallel, such as both time and aggregate
size for detachments to begin. Also, complex cellular behavior can be deduced
from movement, or rather lack thereof, which was interpreted as re-attachment of
\INA daughter cells to the \ac{hMSC} monolayer. This allowed for measuring the
duration of \nMAina existing until re-attaching and turning into \MAina. This
information was helpful when designing experiments to prove that dissemination
is initiated by cell division, requiring precise timing to capture the detached
daughter cells right after cell division. Together, live cell imaging enabled
key mechanistic insights in understanding the dynamics involved in multicellular
interactions by integrating the study of multiple phenomena at once.

\textbf{Difficulties Connecting Observation with Acedemic Terminology} Exploring video data
begins with the search of scientific novelties. In order to correctly identify
cellular phenomena relevant to the research question, a deep
understanding of cell biology is required, e.g. in field of cell dynamics to
read migratory behavior \cite{nalbantExploratoryCellDynamics2018}. This is a
challenge for both students and experienced researchers, since finding the
academically correct terms to describe observations is difficult, especially for
novel phenomena or a sequence of events that can overlap. After all, cell
biology is taught using textbooks, not videos. For this project in particular,
the used terminology was revised frequently, being caused by the constant
struggle of finding the middle-ground between the precice description of
observations, the compatibility with results from other experiments,
comprehensability, and memorability. Ultimately, comprehensability and
memorability were prioritized to maximize adoption of the new terminology by
other researchers. For instance, \emph{non MSC adherence} was chosen over
\emph{mobile interaction}, \emph{aggregation} over \emph{homotypic interaction},
and \emph{detachment event} over \emph{in vitro metastasis}. In general, the gap
between observations and their description remains a challenge in exploratory
cell biology that might be overlooked. This gap could be bridged by currently
available multimodal \acp{LLM} like \textit{ChatGPT-4o}: These models could
match recorded phenomena with descriptions and images that were amassed in the
literature over decades. By doing so, researchers not only use established
terminology instead of inventing new terms, but also minimize the risk of
missing potential discoveries.

\textbf{Why Hide Videos Behind a Download Link?} A major challenge remains in
how to effectively present these dynamic observations in a publishable format,
as traditional scientific publications and websites are not equipped to display
video data. Instead, it is common practice to assemble video frames into static
figures, presumably to support both online and printed reading habits
\cite{perasDigitalPaperReading2023}. Representative example videos are then
relegated to supplementary data, similar to the publication from Chapter\,1
\cite{kuricModelingMyelomaDissemination2024}. Although supplementary data is
downloaded often, most biomedical researchers favor a presentation of additional
figures and tables directly on the journal's website
\cite{priceRoleSupplementaryMaterial2018}. Given the increasing availability of
video data%
\footdagger{The total number of \texttt{PubMed} articles with \emph{``live cell
        imaging''} has doubled from 1012 in the year 2011 to 1967 in 2023}. %
Embedding video content next to figures and tables on the article's website does
make a compelling case. In fact, the journal \emph{Nature} does offer this
feature already, but rarely used \cite{NatureVideoContent}. In the end, there is
no reason to not present videos alongside figures and tables, as they can be as
informative, and potentially more so. Such new standards can benefit other
fields of medicine, as videos provide the best medium for first aid, medical
emergency and education \cite{guptaDatasetMedicalInstructional2023}.

\textbf{Key Points:} Overall, Live-cell imaging has proven indispensable in
exploratory cell biology, uncovering dynamic cellular phenomena that static
analyses often miss, probably. This is exemplified in this work, where live-cell
imaging shifted the research focus by revealing unexpected cell behaviors, like
detachment during division, emphasizing the need for integrating real-time
observations with molecular data. By making such dynamic processes visible,
live-cell imaging not only enriches our understanding but also challenges us to
enhance how scientific findings are presented, advocating for greater
accessibility of video data in scientific publications.

\newpage


% ======================================================================
% ======================================================================
\unnsubsection{Potential and Challenges of Image Cytometry}%
\label{sec:discussion_potential_microscopy}%
Quantifying microscopy data is critical for both analytic and exploratory
approaches to microscopy: For instance, microscopic assessment of live/dead
cells should produce bar charts presenting cell viabilities
\cite{spaepenDigitalImageProcessing2011}, whereas describing novel phenomena
should be supported by charts proving the reproducibility of claimed
observations. Microscopy data is source of vast amount and types of information:
cell morphology; organelle count, shape, and distribution; membrane and lipid
distribution; protein localization, DNA content, et cetera. However, leveraging
this information has always been limited by the ability to extract quantitative
data from microscopy images \cite{galbraithPumpingVolume2023}. This extraction
process is the essence of \emph{image cytometry}, a field that has seen significant
advances by integrating machine learning for automating image analysis tasks.
\cite{guptaDeepLearningImage2019}. The following sections discuss the
experiences gained from this project in quantifying microscopy data and outlines
potentials and challenges of image cytometry.





\textbf{Considering Automated Analysis for Future Live-Cell Imaging:}
This work would have benefited from computational automation for the analysis of
live-cell imaging, for example, the task of associating \INA cell detachment
with \INA aggregate size and time: Manual analysis consisted of zooming in
closely and watching the time-lapse over and over again until a detachment event
was found. A very tedious and error-prone task that had to be repeated approx.
50 times for every one of four independent videos. Instead of manually counting
the number of single \INA cells across time, a pixel segmentation algorithm
could have been trained to detect cells and background. Single cells would be
discernable from aggregates by filtering cells by size. The count of single
cells would then be representative of detached cells, given that the vast
majority of INA6 cells were part of aggregates.



The workload of manual video analysis motivated the purchase of
\texttt{Intellesis}, a software package by \textit{Zeiss} for the \texttt{Zen}
microscopy software ecosystem. \texttt{Intellesis} is a machine learning-based
pixel segmentation software \cite{ZeissOADFeature}. As a feature
extractor\footterm{\footimagefeatures}{\label{foot:image_features}}, it uses the first
convolution layers of VGG19, which is convolutional neural
network\footterm{\footcnn}{\label{foot:cnn}} \cite{simonyanVeryDeepConvolutional2015}.
\texttt{Intellesis} does not contain a deep neural network for segmentation, but
instead classifies pixel features using a \emph{random forest classifier}.
Random forest is a machine learning algorithm that \dashed{for small
    sets of training images} performs almost as well as deep neural networks, but
are computationally far less demanding \cite{breimanRandomForests2001,
    richardsonDenseNeuralNetwork2023}. A comparable hybrid approach was also used by
\citet{qamarHybridCNNRandomForest2023} to segment images of bacterial spores
into eight distinct pixel classes using only 50 training images. Also, free
alternatives to \texttt{Intellesis} exist, such as Ilastik
\cite{bergIlastikInteractiveMachine2019}.

\texttt{Intellesis} proved useful for segmenting single multi-channel images.
However, live cell imaging adds another layer of complexity to image analysis:
The addition of a time axis encodes the motion of objects and other image
features. This concept can be described with the term \emph{optical flow}
\cite{niehorsterOpticFlowHistory2021}. Mathematically speaking, optical flow is
a vector field that describes the motion of image
features\footref{foot:image_features} between consecutive frames of a video. It
can be used to train machine learning models on video data efficiently
\cite{robitailleSelfsupervisedMachineLearning2022}. Without tricks like optical
flow, machine learning algorithms like \texttt{Intellesis} segment the video
frame by frame, ignoring the feature similarities between frames. This makes
segmentation computationally inefficient, but not impossible
\cite{pylvanainenLivecellImagingDeep2023}.

Together, future analyses of live-cell imaging data could  benefit
from the use of modern machine learning based tools that have been released
recently, as summarised in \citet{pylvanainenLivecellImagingDeep2023}.



% ========
\textbf{Image Cytometry is Precise, Fast, Flexible and Accessible:}
In this study, image cytometry was indispensable for validating prior cell
divisions within the \nMAina cell population by profiling their DNA content. The
complexity of this experiment required a method capable of managing a high
throughput across three subpopulations, four timepoints, and two conditions,
involving up to 24 samples per trial (\apdxref{subapdx:figs}{fig:S3}). Despite
having access to automated \ac{FACS} equipment offered by the Core Unit FACS at
the University of Würzburg, the author saw a more time- and cost-effective
solution in the laboratory microscope equipped with motorized stage top and
\texttt{Intellesis}. This setup scanned 96 different samples in \SI{1.5}{\hour},
and resulting large scans were processed by \texttt{Intellesis} overnight,
quantifying thousands of DNA-stained nuclei. This demonstrated that image
cytometry could match the throughput and precision of \ac{FACS} with modern
standard microscopy equipment (\apdxref{subapdx:figs}{fig:S2}).

The advantages of image cytometry could have of great impact for the future of
cell biology: It is applicable to adherent cell cultures
\cite{roukosCellCycleStaging2015} and provides diverse readouts like structure,
brightness, size, and shape. Moreover, image cytometry's capacity to evaluate
cell viability without the need for staining or expensive analytical chemicals
makes it an exceptionally cost-efficient approach for drug screening, reducing
operational costs to cell culturing and electricity for microscopy
\cite{pattaroneLearningDeepFeatures2021}.  However, challenges such as the need
for sophisticated automation in microscopic scans, including autofocus and
shading adjustments, and the computational demands of AI processing remain.


Interestingly, the author's initial unfamiliarity with image cytometry and
limited experience in image processing did not prevent the effective use of this
technology. This underscores the accessibility of current imaging tools to
biologists without specialized training in image analysis. As confirmed by
recent advancements \cite{nittaRapidHighthroughputCell2023}, image cytometry is
becoming increasingly competitive with established techniques like \ac{FACS}.
Despite its limitations, the simplicity and efficiency of image cytometry could
be pivotal for its broader acceptance and integration into biological research.
The exclusivity of \texttt{Intellesis} to \textit{Zeiss} microscopes could be a
major hurdle, however there are free alternatives offering potentially comparable
accessibility, such as \texttt{ilastik} \cite{bergIlastikInteractiveMachine2019}.



\textbf{Manual Analysis Remains Robust for Complex and Unique Phenomena:}
Many biologists lack the access to tools like \texttt{Intellesis}, or the
computational expertise to automate analysis of microscopy data, often reverting
to manual analysis. This project also utilized manual strategies for the
detailed characterization of dynamic intercellular interactions such as
attachment, aggregation, detachment, and division. This was very time-consuming
and required a thoughtful categorization strategy and a disciplined, bias-free
execution. However, some analysis tasks are simply unfeasable for automation.
For example, this work manually counted if two \INA cells interacted
homotypically due to coming into contact with each other, or by staying
connected as two daughter cells after cell division. Automating such a task
would require a very sophisticated algorithm and developing such would be
unfeasable for a task that unique. Hence, manual analysis is unmatched in terms
of flexibility and complexity of categorizations, when compared to computational
techniques of image processing.





\textbf{Key Points:} In summary, image cytometry significantly enhanced this
project by merging the precision of \ac{FACS} with the cost-efficiency of modern
microscopy. Utilizing \texttt{Intellesis} simplified complex image analyses,
making advanced cytometric techniques more accessible. Manual analysis of image
data remains essential for unique and complex phenomena. While challenges like
automation and software availability persist, the potential of image cytometry
to advance biomedical research and discovery remains substantial.




% ======================================================================
% ======================================================================
\unnsubsection{Technical Considerations for Automated Microscopy}%
\label{sec:discussion_quantifying_microscopy}%
Live-cell imaging and image cytometry leverage hardware and software automation
to capture large amounts of potentially unknown complexity. Careful
consideration of technical aspects is crucial to ensure successful data
acquisition and analysis. This section discusses additional challenges and
considerations encountered in this project.

\textbf{Acquiring Accurate Image Data:}
In order to capture rare cellular events with a frequency sufficient for
statistical analysis, this study chose high temporal resolution and spatial
depth: We utilized \SI{1}{frame} every \SI{15}{\minute}, suitable for tracking
cell migration \cite{huthSignificantlyImprovedPrecision2010}, but too slow for
intricate movements or intracellular processes. Spatial resolution is a
compromise between detail and the total observed surface area. We favored the
latter to allow the exploration of potentially rare events, and acquired a
\dashed{somewhat arbitrarily} large surface area of up to
\SI{13}{\milli\meter\squared}. Ultimately, we assessed only approx. a quarter of
the acuired surface area, as that was sufficient to gather enough events for
each time bin. Such extensive automated video acquisition poses high demands on
microscopy equipment, including an incubation setup and motorized stage top. The
total size of video files can also complicate storage, transfer and analysis.
The raw video data from chapter 1 comprises \SI{80}{GB}
\cite{biostudiesBioStudiesEuropeanBioinformatics}; however, far more data was
acquired due to protocol optimizations and treatments not shown in this work.
File size could have been reduced by acquiring in an 8-bit image format,
although a larger bit-depth could be necessary for precise and/or sensitive
fluorescence microscopy. Minimizing the acquired surface area could have reduced
file size as well, however the meniskus of the medium led to significant shading
effects that complicated the choice of the surface area for phase contrasting.
Also, archiving large surface scans allows for the search of very rare events in
the course of future projects. After all, HDD space is cheap, while re-acquiring
data is not. Hence, exploratory live cell imaging benefits from settings that
are higher-than-required, if raw data is properly documented and remain
accessible.



\textbf{Generating Training Datasets:}
In this project, considerable effort was dedicated to training the machine
learning software \textit{Intellesis} for image segmentation, particularly for
fluorescent images. It was also utilized for phase contrast, yet training
required far more effort in generating annotated training images. Phase contrast
or brightfield images often display low contrast between cell edges and the
background, complicating the task of differentiating individual cells from their
surroundings. Such complexity necessitates extensive annotation of training
images --\,a process that can be both time-consuming and demanding.

To address these challenges and enhance the efficacy of \textit{Intellesis},
pre-processing steps could be incorporated to emphasize essential image features
and reduce irrelevant ones. For instance, edge-enhancing filters can
clarify cell boundaries, or a median filter to suppress noise and
unnecessary details while preserving edges. Especially noise reducing filters
can normalize different sets of images, since machine learning models are
very sensitive to variations in image quality.
These filters, available within the
\texttt{Zen} software suite, help simplify the machine learning task by focusing
the algorithm's learning on pertinent features, thereby potentially reducing the
volume of data needed for effective training.

This approach streamlines the training process for \textit{Intellesis}, enabling
more efficient and accurate segmentation of complex microscopy images. By
refining the feature extraction phase, the project could have improved the
performance of the segmentation algorithm but also significantly cut down on the
labor and frustration typically associated with preparing large sets of
annotated training data.




% ======================================================================
\newpage
\unnsubsection{\textit{\textbf{Conclusion\,2:} Automating Microscopy,
        an Emerging Trend for Exploring Unknown Phenomena?}}%
\label{sec:discussion_conclusion_microscopy}%
This study employed two methods that rely heavily on both hard- and software
automation: live-cell imaging and image cytometry. These methods were
instrumental in investigating the complex interactions between \INA cells and
hMSCs, offering significant insights into myeloma cell behavior. The findings
underscored the critical role of MSC-interactions in shaping cell behavior, for
instance the observed preference of \INA cells for heterotypic interactions with
hMSCs. Live-cell imaging and image cytometry revealed dynamic processes, such as
\INA cell growing into clonal aggregates and subsequent cell detachments. These
processes were pivotal for understanding cryptic RNAseq data, exemplifying how
such techniques complement molecular approaches to understand disease mechanisms
like dissemination.

Live-cell imaging proved instrumental in capturing real-time cellular behaviors
that static methods cannot, such as the detachment of \INA cells during
division. This ability to directly observe dynamic processes provided a deeper
understanding of the cellular mechanisms that may contribute to myeloma
dissemination.  However, it also presented challenges, such as the extensive
manual analysis required, difficulties in identifying observed phenomena, and
the complexities of extracting quantitative data and presenting dynamic
observations in traditional scientific formats. These limitations highlight the
potential benefits of incorporating automation and machine learning in future
research to streamline data analysis \cite{guptaDeepLearningImage2019,
    chengFrontiersDevelopmentLivecell2023}.

Image cytometry facilitated high-throughput and precise analysis of cellular
interactions, despite challenges related to automation and computational
requirements. The integration of manual and automated techniques in this study
facilitated complex analyses, combining the precision of automation with the
flexibility of manual approaches. The accessibility and availability of
live-cell imaging equipment and machine learning software suggest a promising
potential for broader adoption in biomedical research. This could lead to new
discoveries of complex cellular dynamics and multicellular interactions.

These findings underscore a potentially emerging trend of innovative imaging and
analyses techniques which could be driven my recent breakthroughs in machine
learning. This trend could mark a significant advancement over molecular
and static approaches of cell biological research questions.







