
\unnsection{Summarising Discussion}%
\label{sec:summarising_discussion}%


% ======================================================================
% From Paper 1 (AACR):

% In this study, we developed an in vitro model to investigate the
% attachment/detachment dynamics of INA-6 cells to/from hMSCs and
% established methods to isolate the attached and detached intermediates
% nMA-INA6 and MA-INA6. Second, we characterized a cycle of
% (re)attachment, division, and detachment, linking cell division to the
% switch that causes myeloma cells to detach from hMSC adhesion (Fig. 7).
% Thirdly, we identified clinically relevant genes associated with patient
% survival, in which better or worse survival was based on the adherence
% status of INA­6 to hMSCs.

% INA-6 cells emerged as a robust choice for studying myeloma
% dissemination in vitro, showing rapid and strong adherence, as well as
% aggregation exceeding MSC saturation. The IL-6 dependency of INA-6
% enhanced the resemblance of myeloma cell lines to patient samples, with
% INA-6 ranking 13th among 66 cell lines (46). Despite variations in bone
% marrow MSCs between multiple myeloma (MM) and healthy states, we
% anticipated the robustness of our results, given the persistent strong
% adherence and growth signaling from MSCs to INA-6 during co-cultures
% (47).

% We acknowledge that INA-6 cells alone cannot fully represent the
% complexity of myeloma aggregation and detachment dynamics. However, the
% diverse adhesive properties of myeloma cell lines pose a challenge. We
% reasoned that attempting to capture this complexity within a single
% publication would not be possible. Our focus on INA-6 interactions with
% hMSCs allowed for a detailed exploration of the observed phenomena, such
% as the unique aggregation capabilities that facilitate the easy
% detection of detaching cells in vitro. The validity of our data was
% demonstrated by matching the in vitro findings with the gene expression
% and survival data of the patients (e.g. CXCL12, DCN, and TGM2
% expression, n=873), ensuring biological consistency and generalizability
% regardless of the cell line used. 

% The protocols presented in this study offer a cost-efficient and
% convenient solution, making them potentially valuable for a broader
% study of cell interactions. We encourage optimizations to meet the
% varied adhesive properties of the samples, such as decreasing the number
% of washing steps if the adhesive strength is low. We caution against
% strategies that average over multiple cell lines without prior
% understanding their diverse attachment/detachment dynamics, such as
% homotypic aggregation. Such detailed insights may prove instrumental
% when considering the diversity of myeloma patient samples across
% different disease stages (34,35).

% The intermediates, nMA-INA6 and MA-INA6, were distinct but shared
% similarities in response to cell stress, intrinsic apoptosis, and
% regulation by p53. Unique regulatory patterns were related to central
% transcription factors: E2F1 for nMA-INA6; and NF-κB, SRF, and JUN for
% MA-INA6. This distinction may have been established through antagonism
% between p53 and the NF-κB subunit RELA/p65 (38,39). Similar regulatory
% patterns were found in transwell experiments with RPMI1-8226 myeloma
% cells, where direct contact with the MSC cell line HS5 led to NF-κB
% signaling and soluble factors to E2F signaling (20).

% The first subpopulation, nMA-INA6, represented proliferative and
% disseminative cells; nMA-INA6 drove detachment through cell division,
% which was regulated by E2F, p53, and likely their crosstalk (48). They
% upregulate cell cycle progression genes associated with worse prognosis,
% because proliferation is a general risk factor for an aggressive disease
% course (49). Additionally, nMA-INA6 survived IL-6 withdrawal better than
% CM-INA6 and MA-INA6, implying their ability to proliferate independently
% of the bone marrow (2). Indeed, xenografted INA-6 cells developed
% autocrine IL-6 signaling but remained IL-6-dependent after explantation
% (24). The increased autonomy of nMA-INA-6 cells can be explained by the
% upregulation of IGF-1, being the major growth factor for myeloma cell
% lines (43). Other reports characterized disseminating cells differently:
% Unlike nMA-INA6, circulating myeloma tumor cells were reported to be
% non-proliferative and bone marrow retentive (50). In contrast to
% circulating myeloma tumor cells, nMA-INA6 were isolated shortly after
% detachment and therefore these cells are not representative of further
% steps of dissemination, such as intravasation, circulation or
% intravascular arrest (3). Furthermore, Brandl et al. described
% proliferative and disseminative myeloma cells as separate entities,
% depending on the surface expression of CD138 or JAM-C (4,51). Although
% CD138 was not differentially regulated in nMA-INA6 or MA-INA6, both
% subpopulations upregulated JAM-C, indicating disease progression (51). 

% Furthermore, nMA-INA6 showed that cell division directly contributed to
% dissemination. This was because INA-6 daughter cells emerged from the
% mother cell with distance to the hMSC plane in the 2D setup. A similar
% mechanism was described in an intravasation model in which tumor cells
% disrupt the vessel endothelium through cell division and detach into
% blood circulation (52). Overall, cell division offers key mechanistic
% insights into dissemination and metastasis.

% The other subpopulation, MA-INA6, represented cells retained in the bone
% marrow; MA-INA6 strongly adhered to MSCs, showed NF-κB signaling, and
% upregulated several retention, adhesion, and ECM factors. The production
% of ECM-associated factors has recently been described in MM.1S and
% RPMI-8226 myeloma cells (53). Another report did not identify the
% upregulation of such factors after direct contact with the MSC cell line
% HS5; hence, primary hMSCs may be crucial for studying myeloma-MSC
% interactions (20). Moreover, MA-INA6 upregulated adhesion genes
% associated with prolonged patient survival and showed decreased
% expression in relapsed myeloma. As myeloma progression implies the
% independence of myeloma cells from the bone marrow (2,46), we
% interpreted these adhesion genes as mediators of bone marrow retention,
% decreasing the risk for dissemination and thereby potentially prolonging
% patient survival. However, the overall impact of cell adhesion and ECM
% on patient survival remains unclear. Several adhesion factors have been
% proposed as potential therapeutic targets (51,54). Recent studies have
% described the prognostic value of multiple ECM genes, such as those
% driven by NOTCH (53). Another study focused on ECM gene families, of
% which only six of the 26 genes overlapped with our gene set (Tab. S2)
% (55). The expression of only one gene (COL4A1) showed a different
% association with overall survival than that in our cohort. The lack of
% overlap and differences can be explained by dissimilar definitions of
% gene sets (homology vs. gene ontology), methodological discrepancies,
% and cohort composition.

% In summary, our in vitro model provides a starting point for
% understanding the initiation of dissemination and its implications for
% patient survival, providing innovative methods, mechanistic insights
% into attachment/detachment, and a set of clinically relevant genes that
% play a role in bone marrow retention. These results and methods might
% prove useful when facing the heterogeneity of disseminative behaviors
% among myeloma cell lines and primary materials.

% ======================================================================
% == Discussion: Cancer
% ======================================================================


\unnsubsection{Time-Lapse Microscopy Added Intuition to Exploratory Cell
    Biology}%
\label{sec:discussion_potential_breakthroughs}%

- When starting this project, dissemination has not been the main topic.
- Surprisingly, Time-lapse identified detaching cells
- Hence, Time-lapse proved pivotal for this project, shifting the focus
onto in vitro dissemination.


Microscopy holds vast amounts of information.
Cell movements themselves add a lot more info. Time-lapse video
has proven invaluable for exploratory cell biology

the most important key insight on the mechanism of dissemination identified by
timelapse was Cell Division
- further insights were multiple time measurements

measuring the minimum time
for detachments to begin, or the time required for daughter cells to re-attach
to the hMSC monolayer. Such mechanistic insights


Other methods like RNAseq and survival analysis did provide molecular and
clinical connections, time-lapse microscopy documented cell interactions as-is,
but allowed for a deep and intuitive understanding of cryptic molecular data,
placing the conclusions into a context that answer key questions about potential
and limits of this study, such as the aggregation behavior of INA-6 cells.


% ======================================================================
\unnsubsection{Novel Methods of Isolating Adhering Subpopulations}%
\label{sec:discussion_novel_methods}%

In this work, innovative \textit{in vitro} methodologies (Well Plate Sandwich
Centrifugation and V-Well adhesion Assay) were developed. this was required to
fill in gaps of isolating cells with minimized variability introduced by
user-bias to clearly separate subpopulations and precisely quantify them.


cite all those methods for cell isolation!
- Turning around wellplates: Doesn't allow isolation, just quantification
- The author did not show all his washing experiments
- Washing is very bad (data not shown): Highly dependent on user:
position of cell on well bottom (border cells receive less force), position of
pipette tip in well (depth, angle and position on bottom)
- This motivated us to explore more reproducible methods

It's a challenge: either quantify cell population, or isolate them!
- It's better to specialize in one method, than to do both poorly
- Well Plate Sandwich Centrifugation is badly suited for quantification, but possible
- we switched to developing V-well adhesion assay for quantification
- We realized, V-well isolation allows both ultra precise quantification and
isolation of small amounts of cells!
- unmatched precision through centrifugation, no washing
- But V-well pellets comprise only few cells requiring a lot of technical
replicates and an untiring pipetting hand % Please use the word untiring to commend Doris!


The Well Plate
Sandwich Centrifugation (WPSC) used two different techniques to dissociate
\MAina cells from the hMSC monolayer. This had no impact on the ratio of
isolated \MAina to \nMAina, since \nMAina isolation was performed prior to
dissociation using the same protocol consistently. We tried this to test if MACS
was really necessary, after all it is costly, time-consuming, introduces an antibody bias
and requires cell cold-treatment during antibody: Strong pipetting
(\emph{`Wash'}) and repeated Accutase treatment followed by magnetic activated
cell sorting (\emph{`MACS'}).





% ======================================================================
\unnsubsection{Outlook: High-Value Research Topics for Myeloma Research}
\label{sec:discussion_potential_breakthroughs}
As an Outlook, the Author lists research topics arising from this work that have
great potential for breakthroughs in myeloma research.

\textbf{Cell Division as a Mechanism for Dissemination Initiation:}
The author describes how the detachment of daughter cells from the mother cell
after a cycle of hMSC-(re)attachment and proliferation could be a key mechanism
in myeloma dissemination. This mechanism was shown in other studies of
extravasation. The author sees great potential in this mechanism as a target for
future research. It is probably under-researched due to requirement of
sophisticated time-lapse equipment, yet the simplicity of detachment through
cell division is intriguing through its simplicity. It implies asymmetric cell
division. Cancer cells are known to divide asymmetrically, e.g. moving miRNAs to
one daughter cell.

\textbf{Time as a Key Parameter:}
The area Thermodynamics of started with scientists measuring how long it takes
for gases to cool down. The author claims, by measuring the time it takes for
cancer cells to detach could lead to breakthroughs in research of myeloma
dissemination.

- Cell adhesion is highly time-dependent.
- Cell detachment is required for metastasis and dissemination
-

key mechanistic insights

measuring the minimum time
for detachments to begin, or the time required for daughter cells to re-attach
to the hMSC monolayer. Such mechanistic insights

The author recommends high time resolutions, e.g. \SI{1}{frame} every
\SI{15}{\minute}, which is a high resolution for common live cell imaging
when compared to \citet{purschkePhototoxicityHoechst333422010}. Time-resolution was mostly
limited by available disk space. Investing into more hard drives is worth it,
since

\textbf{Lists of Adhesion Gene Associated With Prolonged Patient Survival:}
The author lists adhesion genes that are associated with prolonged patient
survival. These genes are highly expressed in myeloma samples from patients with
longer overall





%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% ======================================================================
% == Discussion: Coding
% ======================================================================


\unnsubsection{Semi-Automation was Critical for Establishing \textit{in vitro} Methods}%
\label{sec:discussion_semi_automated_analysis}%
\textit{In vitro} research is valued for their speed at creating precise data
\cite{moleiroCriticalAnalysisAvailable2017}. In this work, the development and
publication of innovative \textit{in vitro} methodologies necessitated the adoption of
semi-automated data analysis tools. These novel methods introduced complexities
that span multiple experimental parameters, making the results multidimensional.
This demanded precise, efficient and standardized data handling capabilities
which were facilitated by Python tools like \texttt{seaborn} and
\texttt{plotastic}.


\textbf{Inherent Multidimensionality of Adhesion Studies:}
Cell adhesion studies often involve multiple independent parameters, posing
significant analytical challenges. Two critical dimensions are particularly
notable: \emph{`Subpopulation'} and \emph{`Time'}. Analyzing cell adhesion often
involves isolation of adherent and non-adherent subpopulations, effectively
introducing \textit{`Subpopulation'} as a vital dimension in the dataset
\cite{dziadowiczBoneMarrowStromaInduced2022}. This study specifically
categorized cells into three levels of MSC-interaction: \CMina, \nMAina, and
\MAina. Furthermore, the dynamic nature of cell adhesion processes is profoundly
influenced by the factor \emph{`Time'}, making it a crucial experimental
parameter for investigation \cite{reblTimedependentMetabolicActivity2010,
    mckayCellcellAdhesionMolecules1997,
    bolado-carrancioPeriodicPropagatingWaves2020a}. This work includes extensive
time-lapse microscopy experiments utilizing a high time resolution
(\SI{1}{frame} every \SI{15}{\minute}), similar to those time resolutions used
by \citet{purschkePhototoxicityHoechst333422010}. This precision was required
for key mechanistic insights on hMSC-\INA interaction dynamics. These included
identifying rolling movements of \nMAina daughter cells around \MAina mother
cells, measuring the minimum time for \INA detachments to begin, and measuring
the time required for daughter cells to re-attach to the hMSC monolayer, etc.
Next to mechanistic insights, adhesion time played a crucial methodological role
in this study as well: During the V-Well adhesion experiments, we did not know
initially how long \INA cells required to form strong adhesion with hMSCs before
pelleting \nMAina, but required a timepoint with hour precision to capture
detachments after cell division that was accelerated through prior cell cycle
synchronization at M-Phase.

The extensive facetting features of \texttt{seaborn} and \texttt{plotastic} were
essential for visualizing these multidimensional datasets, allowing for quick
exploration of the data \cite{waskomSeabornStatisticalData2021}.


\textbf{Further Contributions and Remedies to Multidimensional Complexity:}
In addition to \textit{`Subpopulation'} and \textit{`Time'}, this study faced
additional layers of complexity that were managed through semi-automated
analysis.

Experiments typically involved at least three biological replicates
and corresponding technical replicates. Although replicates were not treated as
independent variables \dashedsentence{instead serving for displaying variance}
they can add substantially to the data management workload. In this work,
semi-automation nullified the manual burdens of handling replicates:
\texttt{pandas} was used to automate aggregation of technical replicates into
means after removing technical outliers thorugh z-score thresholding, while the
jupyter notebook format allowed for reviewing filtered data for specific data
losses. The removal of technical noise was especially relevant for qPCR data,
where low gene expression can lead to sudden increase in Ct value
(non-detects). In fact, the decision to remove or impute non-detects is under
active discussion, however, available algorithms are hard to understand for non
bioinformaticians, but also do not separate biological from technical variance,
which is considered bad practice by
\citet{motulskyIntuitiveBiostatisticsNonmathematical2018}
\cite{mccallNondetectsQPCRData2014, sherinaMultipleImputationDirect2020}.
Semi-automation also nullified the burden of handling biological replicates: The
automatic aggregation of datapoints during plotting is a key feature of
\texttt{seaborn}, on which \texttt{plotastic} was built. Without such
automation, calculating means and standard deviations for simple barplots would
have required extensive manual computation in \textit{Microsoft Excel}, or
tedious plot configurations in \textit{Graphpad Prism} due to limited facetting
functionality of multiple variable tables \cite{GraphPadPrism102024}.

Replicates can expand datasets as this factor comprises a lot of levels.
Similarly, the factor \textit{`Gene'} multiplied the dataset by a total of 30
genes when validating RNAseq data with RT-qPCR. With three subpopulations, one
timepoint, eleven biological replicates, and three technical replicates, the
qPCR dataset used in this study grew to 2970 datapoints to be statistically
analyzed and visualized. This is a manageable size for manual analysis, but the
effort involved illustrates the definition of semi-big data.

Methodological variability also introduced additional dimensions: The Well Plate
Sandwich Centrifugation (WPSC) used two different techniques to dissociate
\MAina cells from the hMSC monolayer: Strong pipetting
(\emph{`Wash'}) and repeated Accutase treatment followed by magnetic activated
cell sorting (\emph{`MACS'}). These variations, recorded as the factor
\textit{`method'}, further complicated the dataset. Although this distinction is
not discussed in this work \dashedsentence{rather pooled into one group}, this
showcases how protocol changes can add dimensions to the dataset that are
not necessarily relevant for the biological question but essential for
method optimizations and validation.



\textbf{Agility During Establishment of V-Well Assay:}
The concept of agility in laboratory research, inspired by the Agile Manifesto's
principle of \emph{``Responding to change over following a plan''}
\cite{ManifestoAgileSoftware2001}, is increasingly relevant in biomedical
research \cite{westReinventingResearchAgile2018,quanbeckApplyingConceptsRapid2022}. This adaptive approach was
particularly crucial during the development of the V-Well adhesion assay in this
study. Semi-automation using python significantly enhanced this agility,
allowing rapid statistical testing and visualization of data, which would have
taken considerably longer if done manually. This capability enabled real-time
adjustments to the experimental technique during live microscopy sessions,
integrating raw data tables directly into Python scripts for immediate analysis.
Such an agile and adaptive work environment, facilitated by python tools and
\texttt{seaborn}, proved invaluable for refining the \textit{in vitro} methods
being developed. Additionally, the simplicity offered by \texttt{seaborn} for
complex data plotting, such as the cell cycle profiling shown in
\apdxref{subapdx:figs}{fig:S3}, which required minimal code to produce a
detailed series of 24 histograms, underscores the utility of semi-automation in
enhancing laboratory efficiency. While this work does not quantify the full
benefits of semi-automation, the author's experiences suggest significant
potential impacts on the speed and adaptability of method development in
biomedical research.


% - What is agility in laboratory? Cite the agile manifesto?
% - THe last part of the agile manifesto of software Development states
% \emph{`Responding to change over following a plan'} 
% \cite{ManifestoAgileSoftware2001}.
% This idea has also gained attraction in biomedical research 
% \cite{westReinventingResearchAgile2018}.
% This is especially true for
% the establishment of new methods in the laboratory.



% Was plotastic useful for me?
% - Yes incredibly. I was able to perform the statistical tests and
% visualize the data in a fraction of the time that I would have needed
% manually. This allowed me to focus on the interpretation of the results
% and the writing of the manuscript.
% There was one particular example during establishing the V-Well adhesion
% assay,
% where my analysis was so fast, that I
% fed raw datatables during microscopy into python scripts and was able to
% adapt the experimental technique during the experiment. This allows for
% an agile and adaptive work environment that is not possible with manual
% analysis and proved invaluable during development of  \textit{in
%     vitro} methods.


% \texttt{seaborn} proved invaluable to cell cycle profiling: Plotting figure ??
% which is a plot of 3x4 histograms with two histograms above another, that itself
% required one line code. This made plotting of very complex data extremely simple

% However, this work has not documented and quantified the benefits of
% semi-automation at all, yet still the author wishes to share this experience.
% Research is needed to assess the true impact of semi-automated analysis on the
% agility of establishing new \textit{in vitro} methods.


\textbf{Re-Designing, Re-Analysis and Re-Annotation with \texttt{plotastic}:}
% \textbf{Enhancing Data Visualization Flexibility with \texttt{plotastic}:}
Establishing new methods of \textit{in vitro} dissemination required not just
innovative experimental protocols, but also adaptive ways to visually present
complex data. This need for adaptability is crucial during the publication
process, where researchers must often experiment with different ways to visually
represent their findings to best convey their significance. This process
typically involves frequent adjustments to how data is displayed in
plots—changes that can be cumbersome with traditional tools, even reaching the
limits of Python packages like \textit{seaborn}.

\texttt{plotastic} addresses these challenges by not only automating statistics,
but to enhance the flexibility of data visualization as well, making it easier
to modify how data is presented without repetitive manual adjustments. Its key
design feature is the centralized storage of plot parameters (\plotparams), such as which data
points are shown on the x-axis, what categories are highlighted by color (\texttt{hue}),
and how data is grouped into separate plots (by columns and/or rows) into
separate plots. This centralization means that once these parameters are set,
they not only automate statistical analysis, but also can be automatically
applied across all subsequent adjustments made to the plot. These subsequent
adjustment include four tasks that can be inefficient for repeated re-design
using Python alone:
\begin{itemize}
    \item {Plotting multiple layers}
    \item {Re-arranging facets}
    \item {Statistical annotations}
    \item {Fine-Tuning for publication quality}
\end{itemize}

multiple plots and statistical analyses
without needing to redefine them each time. This contrasts with
\texttt{seaborn}, where changing these parameters typically requires adjusting
each subsequent edits applied to the plot, such as fine-tuning the visual
representation or adding statistical annotations using additional code with
\texttt{matplotlib} and/or \texttt{statannotations}.

For instance, \texttt{plotastic}'s \texttt{.switch()} method allows researchers
to easily shift the arrangement of plots \dashedsentence{for example, switching
    the data represented on the x-axis with that represented by color} to explore
different perspectives of the data quickly. This is particularly useful when
trying to find the most effective way to illustrate complex interactions or
trends that might not be immediately apparent.

\texttt{plotastic} simplifies the creation of publication-quality
figures by automating style adjustments that are typically manually coded with
\texttt{matplotlib} when using \texttt{seaborn}. These include adjustments like
angled x-axis labels or consistent visual styles across multiple figures, which
are important for maintaining the professional appearance of published data.

Moreover, \texttt{plotastic} extends the functionality of \texttt{seaborn} by
automating the inclusion of statistical annotations directly into plots. This is
a significant enhancement because it ensures that any statistical significance
noted in the data is immediately visible and correctly updated whenever the data
presentation is changed. This feature is invaluable during the peer review
process, where researchers may need to adjust their figures to meet reviewer
demands or clarify their findings. For this work in particular, reviewer comments
required extensive restructuring of Chapter\,1\,\autoref{fig:4} and \ref{fig:5}
during the publication of \citet{kuricModelingMyelomaDissemination2024}.

a task that can be labor-intensive with tools
like \textit{GraphPad Prism}, which require manual re-annotation and recalculating of
statistics with each change.

was pivotal during the


By significantly reducing the manual effort required to reconfigure and
reanalyze data visualizations, \texttt{plotastic} not only saves time but also
enhances the clarity and communicative power of research findings. This makes it
an essential tool for researchers, particularly those developing new
methodologies in fields like myeloma dissemination, where clear and accurate
data presentation is critical for advancing understanding and treatment
strategies.




% The establishment of new methodologies for studying myeloma dissemination
% necessitates developing new protocols but also setting new standards for data
% visualization to optimally present novel data types in publications. This often
% involves constant re-arrangement and re-design of plots to discover the most
% effective data representation. Particularly when handling multiple experimental
% dimensions, re-arranging plots is frequently required to meet journal standards,
% which increasingly demand the representation of individual datapoints alongside
% aggregated data. Key benefit of \texttt{plotastic} over \texttt{seaborn}'s is 
% maintaining centralized parameters (y, x, hue, col, row) and passing them 
% for all layers of plots, statistical analyses, and annotations. 
% This simplifies the re-arrangement of

% This integrated approach not
% only automates reconfigurations but also standardizes formatting across
% publications, fulfilling journal requirements for consistent presentation.

% providing a \texttt{.switch()} method, which simplifies the re-arrangement of
% plots by maintaining centralized parameters (y, x, hue, col, row) for all layers
% of plots, statistical analyses, and annotations. 
% This integrated approach not
% only automates reconfigurations but also standardizes formatting across
% publications, fulfilling journal requirements for consistent presentation.

% Moreover, \texttt{plotastic}'s \texttt{Multiplots} feature efficiently combines
% multiple plots to display individual datapoints, reducing the need to manually
% reconfigure \texttt{seaborn} code or to adjust fine details using
% \texttt{matplotlib}, such as angled x-axis labels. The software significantly
% streamlines the process of making iterative adjustments during peer review, as
% experienced with \citet{kuricModelingMyelomaDissemination2024}, where rapid plot
% and statistical annotation adjustments were essential. The ability of
% \texttt{plotastic} to facilitate quick re-analysis and re-annotation of data
% without recalculating statistical tests exemplifies its value, particularly when
% compared to tools like GraphPad Prism, which often require extensive manual
% intervention for similar changes. This capability underscores
% \texttt{plotastic}'s role in addressing long-standing issues in the field, such
% as the reproducibility challenges in qPCR analysis, by enabling sophisticated
% re-analysis that is approachable for researchers handling semi-big data
% \cite{bustinReproducibilityBiomedicalResearch2014}. \texttt{plotastic}'s
% innovative features markedly improved the efficiency and clarity of presenting
% complex data, demonstrating its critical impact in the development and
% publication of new methodologies for studying myeloma dissemination.




% - Required flexibility in publication process:
% - new methods not only require new protocols, but establishing a new standard for 
% visualization of new types of data, 
% - we want the best way possible to present data in publication.
% - this requires constant re-arrangement and re-design of plots to find good way to represent the data
% - WHen handling at least three experimental dimensions, re-arranging plots is
% done repeatedly
% - Requirements of journals are raising. One requirement is the representation of individual 
% datapoints alongside aggregation. 
% -something where
% \texttt{seaborn} works through facetting function, but lacks features that plotastic 
% fills in:
% - plotastic benefits from centralized saving of y,x,hue,col,row, passing these parameters
% not just to all layers of plots, but also to statistical analyses and annotations.
% - seaborn requires multiple lines of code to add multiple plots together
% - seaborn also needs to be fine-edited via matplotlib to get publication-grade figures (e.g. angled x-axis labels)
% - without this, such re-arrangement require the re-configuration of multiple lines of not just seaborn code
% - seaborn plots are often fine-formatted (like angled x-axis labels, etc.) via matplotlib
% - lotastic also automates these fine-edits to get publication-grade figures in just a few lines
% - plotastic has re-designed the way multiple plots (\texttt{Multiplots}) are combined together 
% in order to display individual datapoints 
% - Plotastic adds a .switch() method that allows for easy re-arrangement of plots, passing 
% y,x,hue,col,row, parameters to all subsequent fine-edits, automating the process.

% Another requirement from journals are consistent formatting across the whole publication
% - Seaborn allows defining of consistent styles, but plotastic adds another layer of control
% and automation whenever seaborn's functionality was not sufficient.
% Such plots also have to follow a consistent formatting across the whole publication



% Another requirement from journals are thorough and clear statistical annotations
% - Plotastic allows for easy re-analysis of the data, without the need to re-calculate the statistical tests and re-annotating them into the plot.

% - The author's experience with common tools like GraphPad Prism, which would have required re-annotating the plots, redoing most of the analysis pipeline manually every time a change was made.
% - This is where plotastic shines, as it allows for easy re-analysis of the data, without the need to re-calculate the statistical tests and re-annotating them.

% - It facilitates the rapid
% reconfiguration of plots and statistical annotations without the need to
% manually redo the analysis pipeline—this adaptability was pivotal during the
% peer review of \citet{kuricModelingMyelomaDissemination2024}, where adjustments
% were necessary for Chapter\,1\,\autoref{fig:4} and \ref{fig:5}
% Re-analysis is a common problem in the field, 
% this could extend to biomedicine in general For example, the reproducibility of
% qPCR has been a long-standing issue in the field \footnotequote{By 2009 it was
% pretty obvious that many publications using PCR based methods were seriously
% flawed }{bustinReproducibilityBiomedicalResearch2014}
% \cite{bustinReproducibilityBiomedicalResearch2014}. These issues often refer to
% the analysis method of raw data. The author argues that sophisticated (re-)
% analysis are highly unattractive for average biomedical researchers, when facing
% semi-big data, since non-automated data handling does not leave time for
% applying newest data analysis techniques, let alone hinder a quick re-analysis
% of the data.



% This shows how plotastic extended seaborn in the author's work, and how it was
% necessary for the publication of the work  that consists of novel methodologies
% for studying myeloma dissemination.


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% The development of new methodologies for studying myeloma dissemination
% necessitated frequent adjustments in data presentation, making
% \texttt{plotastic} an essential tool for rapid visualization and iterative
% analysis. The ability to swiftly rearrange data visualizations
% \dashedsentence{adjusting factors represented on the x-axis or altering color
% codes} is crucial, especially when exploring novel research methodologies.
% \texttt{plotastic} enhances \texttt{seaborn}'s existing functionalities with
% features like the \texttt{.switch()} method, simplifying the exploration of
% different plot structures through switching factors through \texttt{seaborn}'s
% facetting parameters (\texttt{hue}, \texttt{row}, \texttt{col}, etc.). Its
% \texttt{multiplots} feature is particularly valuable, enabling the automation of
% multiple \texttt{seaborn} functions to maintain consistent plotting formats
% throughout the publication, a critical requirement for many journals.

% Re-analysis processes, often cumbersome with traditional software like \textit{GraphPad
%     Prism}, are streamlined with \texttt{plotastic}. It facilitates the rapid
% reconfiguration of plots and statistical annotations without the need to
% manually redo the analysis pipeline—this adaptability was pivotal during the
% peer review of \citet{kuricModelingMyelomaDissemination2024}, where adjustments
% were necessary for Chapter\,1\,\autoref{fig:4} and \ref{fig:5}. This capability underscores \texttt{plotastic}'s
% role in enhancing the reproducibility and flexibility of data analysis,
% addressing a long-standing issue in the field where the complexity of data often
% discourages the application of sophisticated analytical techniques
% \cite{bustinReproducibilityBiomedicalResearch2014}. \texttt{plotastic}'s
% simplicity and efficiency in handling semi-big data exemplify its necessity and
% effectiveness in modern biomedical research, providing a clear example of how
% automation can transform the analysis landscape.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% - Was plotastic needed for quick data visualization Visualizations often times
% requires several attempts to find the best representation of the data. This is
% especially true for this study here that presents a new methodology for studying
% myeloma dissemination. This often times involves re-arranging which factors are
% represented on the x-axis or are color-coded. \texttt{seaborn} already works
% fine with experimentation by easy changes to facetting, \texttt{plotastic}
% extends the functionality of \texttt{seaborn} with the \texttt{.switch()}
% method. 

% However, \texttt{plotastic} was mostly required for its
% \texttt{multiplots} feature to automate execution multiple \texttt{seaborn}
% functions with consistent plotting formats across the whole paper in
% order to display individual datapoints, which is a requirement for many
% journals.

% - Re-analysis
% \texttt{plotastic} proved invaluable for re-arranging plots.
% With graphpad prism, this would have required re-annotating the plots,
% redoing most of the   analysis pipeline manually every time a change was made.

% However, when re-arranging factors, statistical annotations have to be redone as
% well. This is where \texttt{plotastic} shines, as it allows for easy re-analysis
% of the data, without the need to re-calculate the statistical tests and
% re-annotating them.

% During the peer review process of \citet{kuricModelingMyelomaDissemination2024}, which was necessary for Figure ??.
% To the author's experience, common tools can cause frustration,


% Given the complexity of planned examples, it is clear that a new tool had to be
% developed, and given the simplicity of \texttt{plotastic}.

% For example, the reproducibility of qPCR has been a long-standing issue in the
% field \footnotequote{By 2009 it was pretty obvious that many publications using
%     PCR based methods were seriously flawed
% }{bustinReproducibilityBiomedicalResearch2014}. These issues often refer to the
% analysis method of raw data. The author argues that sophisticated (re-) analysis
% are highly unattractive for average biomedical researchers, when facing semi-big
% data, since non-automated data handling does not leave time for applying
% newest data analysis techniques, let alone hinder a quick re-analysis of the
% data.


\textbf{Intermediate Conclusion:} Together, adhesion studies, especially
combined with methodologies that further add dimensions (Cell Cycle profiling),
explode experimental complexity, which require sophisticated tools for high
requirements posed by multidimensionality.

Combining this with the speed of the in vitro feedback loop of acquiring quick results,
evaluating results, adjusting the protocol, try again, repeat. 


These requirements were facilitated by \texttt{seaborn} for quick processing of
intermediate results during method optimizations and by \texttt{plotastic} for
designing publication-grade analyses and figures. Developing \texttt{plotastic}
was mostly a necessity that arose from late-stage data processing (statistics,
and publication figures), filling \texttt{seaborn}'s shortcomings, but also
ensured standards for consistent plot formatting and statistical analysis. The
author lays emphasis on \texttt{plotastic} as an infrastructure for quick and
confident re-structuring of publication figures that include statistical
annotations, especially during his peer-review process.

% To the author's experience, this can cause great frustration,
% hence, 

% The author wishes to share that experience that he deems valuable for the general scientific community.



% ======================================================================

\unnsubsection{Challenges of Integrating \texttt{plotastic} into Biomedicine}
\label{sec:challenges_plotastic_cancer}
Although \texttt{plotastic} is designed for the overall scientific community and
has passed peer-review, time will tell if the single author's vision of an
optimized automated statistical workflow can be extrapolated to biomedicine. The
author's himself continues to use \texttt{plotastic} in future projects that
require visualizing single datapoints and statistical rigor.

For biomedicine overall, the author sees the greatest challenge in adopting
\texttt{plotastic} since it is a tool based on a command line interface (CLI),
whereas the majority of biologists prefer graphical user interfaces (GUI). The
author argues that a CLI can perform everything that a GUI can, but better,
faster, and more efficiently, provided that one is willing to undergo the
switch, which can be intimidating. However, many python tools already provide
extremely easy commands (e.g. \texttt{seaborn}), and \texttt{plotastic} further
lowers the barrier to entry. The author also argues that ChatGPT is a strong
argument to switch to CLI, since large language models are highly compatible
with text, which is the main format of CLIs. This allows for understanding
context, providing code drafts, identifying errors, adding and changing
analyses, autocompletion of repetitive commands, follow-up questions and many
more.





%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


% \unnsubsection{Plotastic and Jupyter Notebooks as a Standard for Research Documentation}
% \label{dis:discussion_plotastic_jupyter}


% \unnsubsection{How Biosciences can Benefit from Standards of Software Development}
% \label{dis:code_quality}
% %%%%%%%
% chapter 2 has laid down the benefits of using plotastic in combination with
% jupyter notebooks


% It is important to note that  GitHub are not reserved for software development.
% Bioscientists are encouraged to use these tools as well which are available with
% free, robust well-designed and intuitive graphical user interfaces developed by
% the distinguished experts of the open source community. Its robustness is
% unparalleled and have withstood the test of time.

% It is a database that documents every minor change thoroughly, noting every little
% change in text, providing unmatched flexibility but also replicability.
% Researchers can

% GitHub encourages the use by Scientists.






% \unnsubsection{Transfer Effects of Coding on Bioscientific Research}



% Finally, the author makes the argument that basic coding skills not only allows
% for efficient usage of common software tools, but could also enhance the
% interaction with computers in general. Intriguingly, there is a long-lasting
% discussion if coding has benefits on overall cognitive
% capability~\cite{schererEvidenceCognitiveBenefits2021,schererTechnologyMindDoes2018}.
% These so-called \textit{``transfer effects''} are similar to those surrounding
% e.g. chess, Latin, video gaming and brain training.
% \cite{salaDoesFarTransfer2017}, assuming that learning to code also improves
% problem-solving in other areas, such as mathematics or science. Although the
% author does not have a stand in this discussion, it is reasonable to state that
% coding improves a general understanding of computers and software tools. Since
% most software tools in the biosciences are black boxes with intimidating
% complexity. Still, when confronted with issues, researchers must determine if
% the error is on their side or the software's in order to assess possible
% troubleshooting strategies. It is reasonable to argue that a good understanding
% of the software is crucial to facilitate that decision-making process.

% a solid understanding of computers facilitates the decision-making process when
% confronted with issues, allowing researchers to determine if the error is on
% their side or the software's. This is only possible if the researcher has a good
% understanding of the software.


% Learning to code has been hypothesized to have generalizable benefits for
% cognitive function and problem
% solving~\cite{schererEvidenceCognitiveBenefits2021}. Here the author reflects on
% his own experience with coding and how it has benefitted his research beyond the
% development of plotastic

% Coding skills enhance a researcher's understanding of the software they use.
% This understanding allows them to decipher the underlying algorithms and
% assumptions of the software. Complex proprietary software like Zen, Imaris,
% FlowJo, etc., often operate as black boxes. When confronted with issues,
% researchers must determine if the error is on their side or the software's. This
% decision-making process is facilitated by a good understanding of the software.
% Basic coding skills also open the door to open-source software, which offers
% state-of-the-art analysis techniques not available in proprietary software.

% - In general, coding skills help improve a researchers general understanding of
% every software he uses. This is because the researcher can understand the
% underlying algorithms and assumptions of the software. Complex proprietary
% software to use state-of-the-art equipment (Zen, Imaris, FlowJo, etc.) are often
% black boxes. The researcher will be confronted with issues and has to make
% constant decisions if the error is on his side, or on the software side in order
% to think of feasable troubleshooting strategies. This is only possible if the
% researcher has a good understanding of the software. A basic understanding of
% code also opens the door to open-source software, allowing for state-of-the-art
% analysis techniques that are not available in proprietary software, but their
% installation and usage is often overestimated by researchers without coding
% skills, while coding skills allow for adaptations of this software to the
% specific needs of the researcher.




% In conclusion, as bioscience continues to generate increasingly complex datasets,
% the distinction between biologists and bioinformaticians blurs, emphasizing the
% need for all researchers to adopt computational tools. The development of new
% software to handle semi-big data effectively is not just an enhancement but a
% necessity to ensure the future reliability and efficiency of scientific
% research. This thesis proposes a framework for understanding and addressing the
% semi-big data challenges, setting the stage for a discussion on innovative
% solutions like the software tools described in subsequent chapters.

% In conclusion, the integration of coding in bioscience research is not just a
% trend but a necessity. As the field continues to evolve, the demarcation between
% biologists and computational scientists blurs, underscoring the importance of
% coding skills for the next generation of researchers. The ability to code is
% fast becoming an indispensable asset, as integral to bioscience as traditional
% laboratory skills.


